---
title: "Homework 6"
author: "Lu Qiu"
date: "2023-11-29"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%",
	dpi = 200
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1
Create a `city_state` variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom `victim_race` is white or black. Be sure that `victim_age` is numeric.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New York. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race.

### Problem 2
For this problem, we’ll use the Central Park weather data similar to data we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

set.seed(1)
```

The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:

*  \( r^2 \)
* log(β̂ 1∗β̂ 2)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1)
. Note: broom::glance() is helpful for extracting r̂ 2
 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 1∗β̂ 2)
.

First, generate bootstrap samples.
```{r}
boot_sample = function(df){
  
  sample_frac(df, replace = TRUE)
}

boot_straps =
  tibble(strap_number = 1:5000) |>
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(weather_df))
  )
```


Next, fit the linear model with tmax as the response and tmin and prcp as predictors.
```{r}
boot_results =
  boot_straps |>
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    r_squared = map(models, ~ broom::glance(.)$r.squared),
    beta_product = map(models, ~ log(broom::tidy(.)$estimate[2] * broom::tidy(.)$estimate[3]))
  ) |>
  select(strap_number, r_squared, beta_product) |>
  mutate(beta_product = unlist(beta_product),
         r_squared = unlist(r_squared))
  
```

Plot the distribution of estimates.
```{r}
# Plotting distribution of r_squared
boot_results |>
  ggplot(aes(x = r_squared)) + 
  geom_density() +
  labs(title = "Distribution of R-squared Estimates")

# Plotting distribution of log(beta1 * beta2)
boot_results |>
  ggplot(aes(x = beta_product)) +
  geom_density() +
  labs(title = "Distribution of log(beta1 * beta2) Estimates")
```

The r-squared value is high and appears to follow a normal distribution, centered around ~0.92. However, the distribution of log(beta1 * beta2)  is somewhat left-skewed.
 
Calculate 95% confidence intervals
```{r}
# CI for r_square
boot_results |>
  summarize(
    ci_lower = quantile(r_squared, 0.025),
    ci_upper = quantile(r_squared, 0.975)) |>
  knitr::kable(digits = 3)
  

# CI for beta_product
boot_results |>
  filter(!is.nan(beta_product)) |>
  summarize(
    ci_lower = quantile(r_squared, 0.025),
    ci_upper = quantile(r_squared, 0.975)) |>
  knitr::kable(digits = 3)

```


### Problem 3